%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Course Supplement
% LaTeX Template
% By: Ryan Grove
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paper=a4, fontsize=11pt]{scrartcl} % A4 paper and 11pt font size

\usepackage[T1]{fontenc} % Use 8-bit encoding that has 256 glyphs
\usepackage{fourier} % Use the Adobe Utopia font for the document - comment this line to return to the LaTeX default
\usepackage[english]{babel} % English language/hyphenation
\usepackage{amsmath,amsfonts,amsthm} % Math packages

\usepackage{lipsum} % Used for inserting dummy 'Lorem ipsum' text into the template

\usepackage{sectsty} % Allows customizing section commands
\allsectionsfont{\centering \normalfont\scshape} % Make all sections centered, the default font and small caps

\usepackage{fancyhdr} % Custom headers and footers
\pagestyle{fancyplain} % Makes all pages in the document conform to the custom headers and footers
\fancyhead{} % No page header - if you want one, create it in the same way as the footers below
\fancyfoot[L]{} % Empty left footer
\fancyfoot[C]{} % Empty center footer
%\fancyfoot[R]{\thepage} % Page numbering for right footer
\renewcommand{\headrulewidth}{0pt} % Remove header underlines
\renewcommand{\footrulewidth}{0pt} % Remove footer underlines
\setlength{\headheight}{13.6pt} % Customize the height of the header

\numberwithin{equation}{section} % Number equations within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{figure}{section} % Number figures within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)
\numberwithin{table}{section} % Number tables within sections (i.e. 1.1, 1.2, 2.1, 2.2 instead of 1, 2, 3, 4)

\setlength\parindent{0pt} % Removes all indentation from paragraphs - comment this line for an assignment with lots of text

%----------------------------------------------------------------------------------------
%	TITLE SECTION
%----------------------------------------------------------------------------------------

\newcommand{\horrule}[1]{\rule{\linewidth}{#1}} % Create horizontal rule command with 1 argument of height

\title{	
\normalfont \normalsize 
\textsc{Ryan Grove, Clemson University, MATH1080 - 9} \\ [25pt] % Your name, university, class
\horrule{0.5pt} \\[0.4cm] % Thin top horizontal rule
\huge Supplement to MATH1080 General Course Policies and Calender \\ % The assignment title
\horrule{2pt} \\[0.5cm] % Thick bottom horizontal rule
}

\author{Statement:} 

\date{\normalsize "This document is a supplement to the Math
1080 General Course Policies.  If there is a conflict between anything in 
this supplement and the general course syllabus, the general course 
syllabus takes precedence."} 

\begin{document}

\maketitle % Print the title

%----------------------------------------------------------------------------------------
%	Course Information
%----------------------------------------------------------------------------------------

\section*{\textbf{Course Information}}

\textbf{Class Time:} 12:20-1:10pm MWF; 12:30-1:20pm T \\
\textbf{Class Location:} Martin M202 \\
\textbf{Instructor:} Ryan Grove \\
\textbf{Office:} Martin Hall M-306 \\
\textbf{Email:} rgrove@clemson.edu \\
\textbf{Website:} http://people.clemson.edu/~rgrove/ \\
\textbf{SI Leader:} AJ Miller 

%---------------------------------------------------------------------------------------
%	Office Hours
%----------------------------------------------------------------------------------------

\section*{\textbf{Office Hours}}

\textbf{Note:} Office hours in M-306 unless otherwise specified
\begin{itemize}
\item 12:00-12:20pm MWF, 
\item 12:00-12:30pm T, 
\item 1:20-2:00pm T, 
\item 2:00-3:00pm T (in the help center located in Martin M-307)
\item By appointment 
\end{itemize}

%---------------------------------------------------------------------------------------
%	Attendance
%----------------------------------------------------------------------------------------

\section*{\textbf{Attendance:}}

Attendance will be kept track of, but is not mandatory. A huge part of the transition into your professional careers is being where you are supposed to be when you are supposed to be there. If the instructor is ever more than 15 minutes late for class, class is understood to be cancelled for the day.

%---------------------------------------------------------------------------------------
%	Cell Phone/Laptop Usage
%----------------------------------------------------------------------------------------

\section*{\textbf{Cell Phone/Laptop Usage:}}

Students are permitted to use cell phones or laptop computers during lecture or activity time unless otherwise specified by your instructor. If you must answer a call, please excuse yourself from the classroom.

%---------------------------------------------------------------------------------------
%	Section Grade
%----------------------------------------------------------------------------------------

\section*{\textbf{Section Grade:}}

An average of two graded written assignments will be given each week. Graded assignments will consist of take-home quizzes and in-class quizzes. Take-home quizzes will always be given on Monday and due Friday, and in-class quizzes will always be on Friday unless otherwise specified by your instructor.\\

Take-home quizzes and in-class quizzes work make up 10\% or your grade. The grade in this category is usually higher than the overall course average because students can get help with all work in this category (as opposed to tests and the final exam).

%---------------------------------------------------------------------------------------
%	In-class Quizzes
%----------------------------------------------------------------------------------------

\section*{\textbf{In-class Quizzes:}}

In-class quizzes will always be on Friday unless otherwise specified by your instructor. They will almost always half questions from the Webassign homework and half test-like problems.

%---------------------------------------------------------------------------------------
%	Take-home Quizzes
%----------------------------------------------------------------------------------------

\section*{\textbf{Take-home Quizzes:}}

You must show all of the work on the take-home quiz to receive credit. Also, every member of the group must sign the take-home quiz to receive credit. It is \textbf{incredibly important} that you all make sure that everyone in your group understands the material covered on the take-home quizzes as well as the Webassign homework. You all are each other's best resources. \textit{Help each other}.

%---------------------------------------------------------------------------------------
%	Policy for Giving/Receiving Help on Quizzes
%----------------------------------------------------------------------------------------

\section*{\textbf{Policy for Giving/Receiving Help on Quizzes:}}

Take-home quizzes are open-book, open-notes, and team-based. Students are encouraged to work together, compare results, and revise as necessary. Working together \textbf{DOES NOT} mean one person does all the work. Rather, all students should meet up, each compute a result individually, and then compare these results and determine which, if any, answer is correct and why.  Make sure everyone understands the material in your group.\\

We will create groups and the person with the best handwriting in each group can write up the take-home quizzes for each group. Every member of the group must sign the take-home quiz to receive credit.\\

In-class quizzes are individual assignments that follow the same standards as tests (closed-book, closednotes, \textbf{NO} collaboration with others) unless otherwise specified by your instructor.

%---------------------------------------------------------------------------------------
%	Important Websites
%----------------------------------------------------------------------------------------

\section*{\textbf{Important Websites:}}

\begin{itemize}
\item \textbf{MATH 1080 Course Webpage:} http://mthsc.clemson.edu/ug/MthSc108/
\item \textbf{Blackboard:} https://bb.clemson.edu/webapps/portal/frameset.jsp
\item \textbf{WebAssign:} http://www.webassign.net/
\end{itemize}

%---------------------------------------------------------------------------------------
%	WebAssign
%----------------------------------------------------------------------------------------

\section*{\textbf{WebAssign:}}
Each student will work online homework using WebAssign. You will receive an  installation packet with your textbook bundle. Keep this packet! It contains your registration code. You will also need a class key when registering. \\

\textbf{Class Key:} clemson 0008 4158

%---------------------------------------------------------------------------------------
%	Extra Credit
%----------------------------------------------------------------------------------------

\section*{\textbf{Extra Credit:}}
Every week an extra credit activity will be given that is somehow related to professional development that will be worth half on the points (rounded up) that you score on that week's in-class quiz. \\

\textbf{Note:} Extra credit is towards the Section Grade only.

%---------------------------------------------------------------------------------------
%	Final Exam
%----------------------------------------------------------------------------------------

\section*{\textbf{Final Exam:}}
Monday, April 25th from 11:30 am to 2:00 pm 

%---------------------------------------------------------------------------------------
%	Checklist
%----------------------------------------------------------------------------------------

\section*{\textbf{Checklist of things to do immediately:}}
\begin{itemize}
\item Purchase the textbook bundle (if you haven't already).
\item Register with WebAssign and begin homework assignment "Review: The Substitution Rule".
\item Study for course-wide u-substitution quiz on Friday, January 8th.

%----------------------------------------------------------------------------------------

\end{document}